---
title: "Week 1 - The Basics"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
```


## Downloading and Installing R

### Requirments:

Make sure that your computer is updated to the latest version of the operating system

If you don't update your computer you could run into installation issues:

You will need to things: 

1) R - the actual language that you use [here](https://www.r-project.org/)

2) RStudio - the development environment, the main tool for R [here](https://www.rstudio.com/products/rstudio/download/)


##Data Types in R:

Data has different types in R, and the datat type influence the operations and functions you can perform on it:


Type | Example | Uses
-----|---------|---------
Logical | TRUE, FALSE | 10 > 8 = TRUE
Numeric | 1, 43.1, 532.37 | 2 + 2 = 4
Character | "Hello", "Goodbye" | 'This is a string'

## Basic Operations in R

Basic operations:
<center>
Operation | Symbol | Example
----------|--------|--------
Addition  |    +   | 2 + 2
Subtraction  |    -   | 3 - 2
Division  |    /   | 4 / 2
Multiplication  |    *   | 2 * 2
Exponentiation  | ^ or ** | 2 ^ 2 or 2**2
Modulus | %% | 4 %% 2
Integer Division | %/% | 5 %/% 2
</center>


Modulus returns the remainder and integer division returns the result of division without the remainder

R follows normal order of operations - remember PEMDAS

####Practice:

below is some space to practice the operations, when you have written your operation click 'Run Code' or 'Start Over'

```{r basic_operations, exercise=TRUE, exercise.lines = 10}

```

####Check Your Understanding:

Use the space above to find the answer to the following question:

```{r quiz_basics}
quiz(
  question("5 %% 2",
    answer("1", correct = TRUE),
    answer("2.5"),
    answer("2"),
    answer("10")
  ),
  question("158 %/% 25",
    answer("6.32"),
    answer("3950"),
    answer("6", correct = TRUE),
    answer("9.256082e+54")
  ),
  question("3 + 10 * (54 %% 3)",
    answer("0", correct = TRUE),
    answer("30"),
    answer("183"),
    answer("Results in Error Message")
  )
  
)
```


##Declaring Variables:

Now that we know how to do some basic operations, we can learn how to store information for later use.

storing the information is simple: using `<-` is the assignment operation, the variable name goes on the left and the value you want to save goes on the right:

` x <- 2 + 2 `

` hello <- "Hello, world!" `

Variable names should be unique. If I repeat a variable name the most recent assingment of that variable will be the one used in proceeding operations where the variable is called. 

Click 'Run Code' to see which value for x is printed!

```{r variables, exercise=TRUE, exercise.lines = 10}
x <- 2

x <- "Hello, world!"

x
```

You can also store variables within variables:
```{r variables_within, exercise=TRUE, exercise.lines = 10}
x <- 2

y <- 6

z <- x + y

z
```

##Questions, Comments and Feedback

Click the link to leave questions or comments and provide feedback for R Bar!

[here](https://docs.google.com/forms/d/e/1FAIpQLScyjSbVxJxHKeR6amIiBdLE_2kFmlS2NQ96Je4rGzPKDgCSFA/viewform?usp=sf_link)
